---
title: "FIPS code matching"
author: "Zeyu Lou"
date: "2024-05-14"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r packages}
library(readxl)
library(openxlsx)
library(foreign)
library(dplyr)
library(stringr)
```

# 0 Data preparation

First, I manually code the state_id in 'MDP_1980_2008.xlsx'

Import df('MDP_1980_2008.xlsx') as dataframe
Import ref('gz_2010_us_050_00_500k.dbf') as the reference list for data linkage

```{r import data}
df <- read_xlsx("../data/MDP_1980_2008.xlsx", sheet=2)
ref <- read.dbf("../data/shp_county_2010/gz_2010_us_050_00_500k.dbf")
```

# 1 Data cleaning
## 1.1 Clean df
```{r clean df}
sum(is.na(df$County))
df.1 <- df %>%
  mutate(County = str_to_lower(County)) %>%
  mutate(County = str_replace_all(County, " ", "")) %>%
  mutate(County = str_replace_all(County, "county", "")) %>%
  mutate(County = str_replace_all(County, "city", "")) %>%
  mutate(County = str_replace_all(County, "parish", ""))
```

We have five NAs in the County column in total, which are Bay St. Louis (city), Richmond (State: VA), and Washington D.C. I located them by hand using Google. However, I assume Bay St. Louis (city) might belong to Hancock County in Mississippi instead of MO as indicated by MDP_1980_2008.xlsx originally. It might be a typo, I guess. Therefore, I will change it to 'MS' and proceed with the linkage.

*For number 107, the county contains three options. Does this indicate that one company has three locations? Should we create three FIPS codes for one company?*

## 1.2 Clean ref

```{r clean ref}
sum(is.na(ref$NAME))
ref.1 <- ref %>%
  mutate(NAME = str_to_lower(NAME)) %>%
  mutate(NAME = str_replace_all(NAME, " ", "")) %>%
  mutate(NAME = str_replace_all(NAME, "county", ""))
```

# 2 Join data

```{r linkage}
# Make 'state_id' as factor with two digit
df.1 <- df.1 %>%
  mutate(state_id = str_pad(as.character(state_id), width = 2, pad = "0")) %>%
  mutate(state_id = as.factor(state_id))

df.1 <- df.1 %>%
  left_join(ref.1 %>% select(STATE, NAME, COUNTY, GEO_ID, lat, lon), by = c("state_id" = "STATE", "County" = "NAME"))

# check NAs
sum(is.na(df.1$COUNTY))
na_rows <- which(is.na(df.1$COUNTY))
na_rows

# Create fips_id
df.1 <- df.1 %>%
  mutate(fips_id = paste0(state_id, COUNTY))

head(df.1$fips_id)

# export the data
write.xlsx(df.1, '../zeyu_data/MDP_1980_2008_updated.xlsx')
```

# 3 Create the map.

```{r map package}
library(sf)
library(tmap)
library(ggplot2)
library(leaflet)
library(maps)
```

```{r map}
shapefile_path <- "../data/shp_county_2010/gz_2010_us_050_00_500k.shp"
geo_data <- st_read(shapefile_path)

# create the map
map("state", fill = FALSE, col = "black", bg = "white", lwd = 1.5)
text(state.center$x, state.center$y, labels = state.abb, cex = 0.7)

for (i in 1:nrow(df.1)) {
  if (df.1$Winner[i] == 1) {
    points(df.1$lon[i], df.1$lat[i], col = "blue", pch = 1, cex = 2, lwd = 2)
  } else {
    symbols(df.1$lon[i], df.1$lat[i], circles = rep(1, nrow(df.1))[i], inches = 0.1,
            add = TRUE, fg = "red", lty = 2)
  }
}
```



