---
title: "Ultimate output data"
author: "Zeyu Lou"
date: "2024-05-17"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 0. Data preparation

## 0.1 Loading package

```{r packages}
library('haven')
library('xml2')
library('foreign')
library(tidyverse)
library(tidycensus)
library(ipumsr)
library(censusapi)
library(lehdr)
library(tidyUSDA)
library(flextable)
library(tigris)
library(sf)
library(dplyr)
library(stringr)
library(Hmisc)
```

## 0.2 Scan the data

```{r loading data}
example_data <- read_dta("../data/example_data.dta")
# xml_file <- read_xml("../data/BBS_data/North American Breeding Bird Survey Dataset (1966-2022).xml")
routes <- read.csv("../data/BBS_data/routes.csv")
# Migrant <- read.csv("../data/BBS_data/MigrantNonBreeder/Migrants.csv")
MigrantSummary <- read.csv("../data/BBS_data/MigrantNonBreeder/MigrantSummary.csv")
VehicleData <- read.csv("../data/BBS_data/VehicleData.csv")
weather <- read.csv("../data/BBS_data/weather.csv")
ref <- read.dbf("../data/shp_county_2010/gz_2010_us_050_00_500k.dbf")
```

```{r api}
# census_api_key("457e74535811f68ac33753e2d9107cb373e935d4", install = TRUE)
```

# 1. MigrantNonBreeder Data

## 1.1 routes_us

```{r routes_us}
# CountryNum: 124=Canada, 840=United_States, so we filtered out the Canada in our data and create new df routes_us.
routes_us <- routes %>%
  filter(CountryNum == 840) %>%
  mutate(
    StateNum_str = str_pad(StateNum, width = 2, pad = "0"),
    Route_str = str_pad(Route, width = 3, pad = "0"),
    StateNum_Route = paste0(StateNum_str, Route_str)
  ) %>%
  select(-StateNum_str, -Route_str)

counties <- counties(cb = TRUE, resolution = "5m")
counties <- st_transform(counties, crs = 4326)

# Convert Latitude and Longitude to sf objects
points <- st_as_sf(routes_us, coords = c("Longitude", "Latitude"), crs = 4326)
# Perform spatial join to find the county where each point is located
joined <- st_join(points, counties, join = st_within)
# Extract STATEFP and COUNTYFP and merge them into FIPS code
routes_us <- routes_us %>%
  mutate(fips = joined$GEOID)
routes_us

rm(joined)
rm(points)
```

## 1.2 MigrantSummary_us

We need Abundance and Richness from MigrantSummary_us

- Abundance: total number of individual birds observed (regardless of species)

- Richness: total number of different bird species observed 

```{r MigrantSummary_us}
# Merge StateNum and Route into a single variable called StateNum_Route, so we can link this dataset with other datasets.
MigrantSummary_us <- MigrantSummary %>%
  filter(CountryNum == 840) %>%
  mutate(
    StateNum_str = str_pad(StateNum, width = 2, pad = "0"),
    Route_str = str_pad(Route, width = 3, pad = "0"),
    StateNum_Route = paste0(StateNum_str, Route_str)
  ) %>%
  select(-StateNum_str, -Route_str)

rm(MigrantSummary)

MigrantSummary_us <- MigrantSummary_us %>%
  left_join(routes_us %>% select(StateNum_Route, fips), by = "StateNum_Route")
```

Create a new dataset "county_x_year_panel_data.1" based on MigrantSummary_us

```{r}
county_x_year_panel_data.1 <- MigrantSummary_us %>%
  group_by(fips, Year) %>%
  summarise(
    abundance = sum(StopTotal),
    richness = sum(SpeciesTotal)
  )
```


Then we need to build another df about no_routes based on MigrantSummary_us

- no_routes: Number of surveying routes in the FIPS county

```{r no_routes}
no_routes <- routes_us %>%
  group_by(fips) %>%
  summarise(no_routes = n_distinct(StateNum_Route)) %>%
  ungroup()

county_x_year_panel_data.1 <- county_x_year_panel_data.1 %>%
  left_join(no_routes, by = c("fips" = "fips"))
```


# 2. States

## 2.1 Alabama-1

df: county_x_year_panel_data.01

```{r alabama}
# Loading data
alabama <- read.csv("../data/BBS_data/States/Alabama.csv")

# Find the fips for each route
alabama <- alabama %>%
  mutate(
    StateNum_str = str_pad(StateNum, width = 2, pad = "0"),
    Route_str = str_pad(Route, width = 3, pad = "0"),
    StateNum_Route = paste0(StateNum_str, Route_str)
  ) %>%
  select(-StateNum_str, -Route_str)

alabama <- alabama %>%
  left_join(routes_us %>% select(StateNum_Route, fips), by = "StateNum_Route")

county_x_year_panel_data.01 <- alabama %>%
  group_by(fips, Year) %>%
  summarise(
    abundance = sum(SpeciesTotal),
    richness = n_distinct(AOU),
    no_routes = n_distinct(StateNum_Route)
  )

```

## 2.2 Loop

```{r colorado}
process_file <- function(file_path, routes_us, no_routes) {
  data <- read.csv(file_path)
  
  if (all(data$CountryNum == 840)) {
    data <- data %>%
      mutate(
        StateNum_str = str_pad(StateNum, width = 2, pad = "0"),
        Route_str = str_pad(Route, width = 3, pad = "0"),
        StateNum_Route = paste0(StateNum_str, Route_str)
      ) %>%
      select(-StateNum_str, -Route_str)
    
    data <- data %>%
      left_join(routes_us %>% select(StateNum_Route, fips), by = "StateNum_Route")
    
    result <- data %>%
      group_by(fips, Year) %>%
      summarise(
        abundance = sum(SpeciesTotal),
        richness = n_distinct(AOU),
        no_routes = n_distinct(StateNum_Route)
      )
    
    return(result)
  } else {
    return(NULL)
  }
}

files<-list.files("../data/BBS_data/States", full.names = TRUE)

results <- list()

# 遍历每个文件
for (file in files) {
  file_result <- process_file(file, routes_us, no_routes)
  if (!is.null(file_result)) {
    state_name <- tools::file_path_sans_ext(basename(file))
    assign(paste0("county_x_year_panel_data_", state_name), file_result)
    results[[state_name]] <- file_result
  }
}

# 将所有结果合并成一个数据框
county_x_year_panel_data.state <- bind_rows(results)

write.csv(county_x_year_panel_data.state, "../data/county_x_year_panel_data.state.csv", row.names = FALSE)
```

# 3. weather

- duration_minutes: Survey duration (minutes)

  - EndTime - StartTime, and set Null as 0

```{r weather_us}
# Build weather_us data and add fips
weather_us <- weather %>%
  filter(CountryNum == 840) %>%
  mutate(
    StateNum_str = str_pad(StateNum, width = 2, pad = '0'),
    Route_str = str_pad(Route, width = 3, pad = '0'),
    StateNum_Route = paste0(StateNum_str, Route_str)
  ) %>%
  select(-StateNum_str, -Route_str) %>%
  left_join(routes_us %>% select(StateNum_Route, fips), by = "StateNum_Route")

# Find the duration_minutes
time_to_minutes <- function(time) {
  hours <- time %/% 100
  minutes <- time %% 100
  return(hours * 60 + minutes)
}

weather_us <- weather_us %>%
  mutate(
    EndTime_minutes = time_to_minutes(EndTime),
    StartTime_minutes = time_to_minutes(StartTime),
    duration_minutes = EndTime_minutes - StartTime_minutes
  ) %>%
  select(-EndTime_minutes, -StartTime_minutes)

# Modify it as weather_us_summary by Year and fips.
weather_us_summary <- weather_us %>%
  group_by(Year, fips) %>%
  summarise(
    duration_minutes = sum(duration_minutes, na.rm = TRUE),
    obsn = sum(ObsN, na.rm = TRUE)
  ) %>%
  ungroup()

# Add weather_us_summary to the main dataset
county_x_year_panel_data <- county_x_year_panel_data.state %>%
  left_join(weather_us_summary, by = c("Year", "fips"))
```

# 4. VehicleData

```{r VehicleData_us}
# Build weather_us data and add fips
VehicleData_us <- VehicleData %>%
  filter(CountryNum == 840) %>%
  mutate(
    StateNum_str = str_pad(StateNum, width = 2, pad = '0'),
    Route_str = str_pad(Route, width = 3, pad = '0'),
    StateNum_Route = paste0(StateNum_str, Route_str)
  ) %>%
  select(-StateNum_str, -Route_str) %>%
  left_join(routes_us %>% select(StateNum_Route, fips), by = "StateNum_Route")

# Sum the car_total and noise_total
VehicleData_us_summary <- VehicleData_us %>%
  mutate(car_total = rowSums(select(., starts_with("Car"))),
         noise_total = rowSums(select(., starts_with("Noise")))
         ) %>%
  group_by(Year, fips) %>%
  summarise(
    car_total = sum(car_total, na.rm = TRUE),
    noise_total = sum(noise_total, na.rm = TRUE)
  )

county_x_year_panel_data <- county_x_year_panel_data %>%
  left_join(VehicleData_us_summary, by = c("Year", "fips")) %>%
  rename(year = Year)
```

# 5. Label and save the data

```{r label}
label(county_x_year_panel_data$fips) <- "GEOID"
label(county_x_year_panel_data$year) <- "Year"
label(county_x_year_panel_data$richness) <- "richness, countyXyear level"
label(county_x_year_panel_data$no_routes) <- "Number of routes in the given fips"
label(county_x_year_panel_data$duration_minutes) <- "Survey duration, in minutes"
label(county_x_year_panel_data$obsn) <- "Number of person conducted the survey"
label(county_x_year_panel_data$car_total) <- "Number of cars passing during the survey"
label(county_x_year_panel_data$noise_total) <- "Number of stops (out of 50) on a survey route that have excessive noise during t"
label(county_x_year_panel_data$abundance) <- "abundance"

write.csv(county_x_year_panel_data, "../data/county_x_year_panel_data.csv", row.names = FALSE)
```

